Supplemental materials for ``An item response theory analysis of the Matrix Reasoning Item Bank (MaRs-IB)''

### 2022-11-18 manuscript/supplement.tex
camera ready docs

The results of the parameter recovery study are summarized Figure \ref{fig:figS04}. For both models, we observed good recovery of the latent ability parameter (Model 5: $r = 0.873$; Model 6: $r = 0.873$). Furthermore, we observed excellent recovery for the item difficulty parameters (Model 5: $r = 0.962$; Model 6: $r = 0.962$). In contrast, we observed only adequate recovery of the item discrimination parameters (Model 5: $r = 0.667$; Model 6: $r = 0.671$). This result is not altogether unexpected given recent research that recommended at least $N=100$ observations per item for accurate item parameter estimation \cite{konig2020optimized}. This result, however, must be interpreted in context. We demonstrate below that the less-than-ideal recovery of item discrimination parameters does not seriously impact the outputs of optimal test assembly.

@@

Finally, we investigated the effects of item parameter misestimation on test assembly. For each simulated dataset, we submitted the ground-truth (simulated) and model-predicted (recovered) item parameters to test assembly under the same constraints used to generate the MaRs-IB short form measures. We then calculated for each test form its test information function (TIF) and IRT test score reliability \cite{kim2010estimation, nicewander2018conditional}. Crucially, we calculated the TIF and reliability for the short forms using the true (i.e. not recovered) item parameter estimates; therefore, we can quantify how much the psychometric properties of test forms are impacted by suboptimal item selection due to parameter estimation noise. The results are summarized in the third and fourth columns of Figure \ref{fig:figS04}. As is expected, TIF was larger for the test forms made using the true item parameters, indicating that estimation noise can lead to suboptimal item selection. Despite this, estimation noise led to only marginal decrements in IRT test reliability. For item parameters estimated using Model 5, parameter noise resulted in an average loss of score reliability of $\Delta \rho = 0.020$ (SD = 0.014). For item parameters estimated using Model 6, parameter noise resulted in an average loss of score reliability of $\Delta \rho = 0.019$ (SD = 0.013).

>>>>

The results of the parameter recovery study are summarized Figure \ref{fig:figS04}. For both models, we observed good recovery of the latent ability parameter (Model 5: $r = 0.873$; Model 6: $r = 0.873$). Furthermore, we observed excellent recovery for the item difficulty parameters (Model 5: $r = 0.962$; Model 6: $r = 0.962$). In contrast, we observed only adequate recovery of the item discrimination parameters (Model 5: $r = 0.667$; Model 6: $r = 0.671$). This result is not altogether unexpected given recent research that recommended at least $N=100$ observations per item for accurate item parameter estimation \citep{konig2020optimized}. This result, however, must be interpreted in context. We demonstrate below that the less-than-ideal recovery of item discrimination parameters does not seriously impact the outputs of optimal test assembly.

@@

Finally, we investigated the effects of item parameter misestimation on test assembly. For each simulated dataset, we submitted the ground-truth (simulated) and model-predicted (recovered) item parameters to test assembly under the same constraints used to generate the MaRs-IB short form measures. We then calculated for each test form its test information function (TIF) and IRT test score reliability \citep{kim2010estimation, nicewander2018conditional}. Crucially, we calculated the TIF and reliability for the short forms using the true (i.e. not recovered) item parameter estimates; therefore, we can quantify how much the psychometric properties of test forms are impacted by suboptimal item selection due to parameter estimation noise. The results are summarized in the third and fourth columns of Figure \ref{fig:figS04}. As is expected, TIF was larger for the test forms made using the true item parameters, indicating that estimation noise can lead to suboptimal item selection. Despite this, estimation noise led to only marginal decrements in IRT test reliability. For item parameters estimated using Model 5, parameter noise resulted in an average loss of score reliability of $\Delta \rho = 0.020$ (SD = 0.014). For item parameters estimated using Model 6, parameter noise resulted in an average loss of score reliability of $\Delta \rho = 0.019$ (SD = 0.013).
